---
title: "sog_model"
output: html_document
date: "2025-09-26"
---

```{r setup, include=FALSE, warning=F, message=F}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nhlscraper)
library(cluster)
library(MASS)
library(caret)
library(Metrics)
library(glmnet)
library(lme4)
library(gt)
load('train_data.RData')
train_data <- as.data.frame(train_data)
```



# Creating Team Schedule Dataframe
```{r}
get_team_schedule_df <- function(team_tricode, date, season = 20252026) {
  date <- as.Date(date)
  sched <- get_team_schedule(team = toupper(team_tricode), season = season) %>%
    dplyr::filter(gameType == 2) %>%
    dplyr::filter(as.Date(gameDate) < date) %>%
    dplyr::select(all_of(c("id", "gameDate", "startTimeUTC", "venueUTCOffset",
                    "awayTeam.id", "awayTeam.abbrev", "awayTeam.score",
                    "homeTeam.id", "homeTeam.abbrev", "homeTeam.score",
                    "gameOutcome.lastPeriodType")))
  return(sched)
}

# Example
sched_sjs <- get_team_schedule_df("SJS", Sys.Date())
sched_sjs
```


# Picking Team
```{r}
teams <- get_teams() %>%
  dplyr::select('id', 'fullName', 'triCode')
#teams


get_team_id <- function(tricode) {
  id <- teams %>%
    dplyr::filter(triCode == toupper(tricode))
  id <- id$id
  if(length(id)>1)
    id <- 68
  return(id)
}
team_id <- get_team_id('edm')
team_id
```



# Creating Game Shot Reports
```{r}
get_game_shot_report <- function(game_id, team_id) {
  pbp <- get_gc_play_by_play(game_id)
  #head(pbp, 50)
  #colnames(pbp)
  if (nrow(pbp) == 0 || !"typeCode" %in% names(pbp)) {
    return(
      tibble(
        total_shot_attempts = NA_real_,
        shots_on_goal_plus_goals = NA_real_,
        blocked_shots = NA_real_,
        missed_shots = NA_real_
      )
    )
  }
  shot_codes <- c(505, 506, 507, 508)
  ## 505-goal, 506-shot on goal, 507-missed shot, 508-blocked shot
  shot_pbp <- pbp %>%
    dplyr::filter(typeCode %in% shot_codes)
  shot_pbp <- shot_pbp %>%
    dplyr::select(any_of(c('typeCode', 'sortOrder', 'periodDescriptor.number', 
           'details.eventOwnerTeamId', 'details.shotType', 'details.homeScore', 'details.awayScore'))) %>%
    dplyr::filter(details.eventOwnerTeamId == team_id) %>%
    dplyr::filter(periodDescriptor.number %in% c(1,2,3,4))
  #print(shot_pbp$details.eventOwnerTeamId)
  #return(shot_pbp)
  shot_report <- shot_pbp %>%
    mutate(
      total_attempt = 1L,
      sog_or_goal = if_else(typeCode %in% c(505, 506), 1L, 0L),
      blocked = if_else(typeCode == 508, 1L, 0L),
      missed = if_else(typeCode == 507, 1L, 0L)
    ) %>%
    summarise(
      total_shot_attempts = sum(total_attempt, na.rm = TRUE),
      shots_on_goal_plus_goals = sum(sog_or_goal, na.rm = TRUE),
      blocked_shots = sum(blocked, na.rm = TRUE),
      missed_shots = sum(missed, na.rm = TRUE),
      .groups = "drop"
    )
  return(shot_report)
}

gsr <- get_game_shot_report(2025020017, 52)
gsr
```

# Creating Home/Away Splits
```{r}
get_away_splits <- function(team_tricode, date){
  team_id <- get_team_id(tricode = toupper(team_tricode))
  #return(team_id)
  
  away_sched <- get_team_schedule_df(team = toupper(team_tricode), date, season = 20252026) %>%
    dplyr::filter(awayTeam.id == team_id) %>%
    dplyr::select(all_of(c("id", "gameDate", "startTimeUTC", "venueUTCOffset",
                    "awayTeam.id", "awayTeam.abbrev", "awayTeam.score",
                    "homeTeam.id", "homeTeam.abbrev", "homeTeam.score",
                    "gameOutcome.lastPeriodType")))
  n_away <- nrow(away_sched)
  
  game_log <- data.frame(matrix(ncol=4, nrow=0))
  cols <- c('total_shot_attempts', 'sog_plus_goals', 'blocked_shots', 'missed_shots')
  colnames(game_log) <- cols
  
  for (i in 1:n_away){
    game_id <- away_sched$id[nrow(away_sched)-(i-1)] 
    #print(game_id)
    
    game_report <- get_game_shot_report(game_id, team_id)
    game_log <- rbind(game_log, game_report)
  }
  
  away_mean_team_log <- game_log %>% summarise(across(where(is.numeric), mean))
  return(away_mean_team_log)
}


get_home_splits <- function(team_tricode, date){
  team_id <- get_team_id(tricode = toupper(team_tricode))
  #return(team_id)
  
  home_sched <- get_team_schedule_df(team = toupper(team_tricode), date, season = 20252026) %>%
    dplyr::filter(homeTeam.id == team_id) %>%
    dplyr::select(all_of(c("id", "gameDate", "startTimeUTC", "venueUTCOffset",
                    "awayTeam.id", "awayTeam.abbrev", "awayTeam.score",
                    "homeTeam.id", "homeTeam.abbrev", "homeTeam.score",
                    "gameOutcome.lastPeriodType")))
  n_home <- nrow(home_sched)
  
  game_log <- data.frame(matrix(ncol=4, nrow=0))
  cols <- c('total_shot_attempts', 'sog_plus_goals', 'blocked_shots', 'missed_shots')
  colnames(game_log) <- cols
  dummy <- 0
  for (i in 1:n_home){
    #print(game_id)
    game_id <- home_sched$id[nrow(home_sched)-(i-1)] 
    game_report <- get_game_shot_report(game_id, team_id)
    game_log <- rbind(game_log, game_report)
  }
  
  home_mean_team_log <- game_log %>% summarise(across(where(is.numeric), mean))
  return(home_mean_team_log)
}

today <- Sys.Date()
get_away_splits('wsh', today)
get_home_splits('wsh', today)
```


# Creating Shot Game Logs for Team
```{r}
get_team_shot_log <- function(team_tricode, n, date){
  sched <- get_team_schedule_df(team_tricode, date, 20252026)
  team_id <- get_team_id(team_tricode)
  game_log <- data.frame(matrix(ncol=4, nrow=0))
  cols <- c('total_shot_attempts', 'sog_plus_goals', 'blocked_shots', 'missed_shots')
  colnames(game_log) <- cols
  
  #print(team_id)
  for (i in 1:n){
    game_id <- sched$id[nrow(sched)-(i-1)] 
    #print(game_id)
    
    game_report <- get_game_shot_report(game_id, team_id)
    game_log <- rbind(game_log, game_report)
  }
  game_log <- game_log %>% summarise(across(where(is.numeric), mean))
  return(game_log)
}

today <- Sys.Date()
l8_team_log <- get_team_shot_log('SJS', 8, today)
l8_team_log

l8_last <- get_team_shot_log('SJS', 8, '2025-10-26')
l8_last
```


# Total Model Predictors Dataframe
```{r}
get_full_home_preds <- function(tricode, date){
  l8_log <- get_team_shot_log(tricode, 8, date)
  l4_log  <- get_team_shot_log(tricode, 4, date)
  ha_split <- get_home_splits(tricode, date)
  
  l8_log <- l8_log %>%
    dplyr::select(total_shot_attempts, shots_on_goal_plus_goals, missed_shots) %>%
    rename(l8_sa = total_shot_attempts, l8_sog = shots_on_goal_plus_goals, l8_miss = missed_shots)
  l4_log <- l4_log %>%
    dplyr::select(total_shot_attempts, shots_on_goal_plus_goals, missed_shots) %>%
    rename(l4_sa = total_shot_attempts, l4_sog = shots_on_goal_plus_goals, l4_miss = missed_shots)
  ha_split <- ha_split %>%
    dplyr::select(total_shot_attempts, shots_on_goal_plus_goals, missed_shots) %>%
    rename(ha_sa = total_shot_attempts, ha_sog = shots_on_goal_plus_goals, ha_miss = missed_shots)
  
  shot_preds <- cbind(l8_log, l4_log, ha_split)
  return(shot_preds)
}
  

get_full_away_preds <- function(tricode, date){
  l8_log <- get_team_shot_log(tricode, 8, date) 
  l4_log  <- get_team_shot_log(tricode, 4, date)
  ha_split    <- get_away_splits(tricode, date)
  
  l8_log <- l8_log %>%
    dplyr::select(total_shot_attempts, shots_on_goal_plus_goals, missed_shots) %>%
    rename(l8_sa = total_shot_attempts, l8_sog = shots_on_goal_plus_goals, l8_miss = missed_shots)
  l4_log <- l4_log %>%
    dplyr::select(total_shot_attempts, shots_on_goal_plus_goals, missed_shots) %>%
    rename(l4_sa = total_shot_attempts, l4_sog = shots_on_goal_plus_goals, l4_miss = missed_shots)
  ha_split <- ha_split %>%
    dplyr::select(total_shot_attempts, shots_on_goal_plus_goals, missed_shots) %>%
    rename(ha_sa = total_shot_attempts, ha_sog = shots_on_goal_plus_goals, ha_miss = missed_shots)
  
  shot_preds <- cbind(l8_log, l4_log, ha_split)
  return(shot_preds)
}

get_full_home_preds('SJS', today)
get_full_home_preds('SJS', '2025-11-03')
```


# Getting Shots Against Splits
```{r}
# Shots-against ALLOWED per game for a team in a given game
get_game_shot_allowed_report <- function(game_id, team_id) {
  pbp <- get_gc_play_by_play(game_id)
  shot_codes <- c(505, 506, 507, 508) # goal, shot on goal, missed, blocked
  shot_pbp <- pbp %>%
    dplyr::filter(typeCode %in% shot_codes) %>%
    dplyr::select(any_of(c('typeCode', 'sortOrder', 'periodDescriptor.number',
           'details.eventOwnerTeamId', 'details.shotType', 'details.homeScore', 'details.awayScore'))) %>%
    dplyr::filter(details.eventOwnerTeamId != team_id) # events by the opponent → allowed by team_id
  shot_report <- shot_pbp %>%
    mutate(
      total_attempt = 1L,
      sog_or_goal = if_else(typeCode %in% c(505, 506), 1L, 0L),
      blocked = if_else(typeCode == 508, 1L, 0L),
      missed = if_else(typeCode == 507, 1L, 0L)
    ) %>%
    summarise(
      total_shot_attempts = sum(total_attempt, na.rm = TRUE),
      shots_on_goal_plus_goals = sum(sog_or_goal, na.rm = TRUE),
      blocked_shots = sum(blocked, na.rm = TRUE),
      missed_shots = sum(missed, na.rm = TRUE),
      .groups = "drop"
    )
  return(shot_report)
}

# Last-N or full-season per-game shots-against allowed log for a team
get_team_shot_allowed_log <- function(team_tricode, n, date) {
  sched <- get_team_schedule_df(team = toupper(team_tricode), date, season = 20252026)
  team_id <- get_team_id(tricode = toupper(team_tricode))
  num_games <- nrow(sched)
  game_log <- data.frame(matrix(ncol = 4, nrow = 0))
  colnames(game_log) <- c('total_shot_attempts', 'shots_on_goal_plus_goals', 'blocked_shots', 'missed_shots')
  for (i in 1:n) {
    game_id <- sched$id[num_games - (i - 1)]
    game_report <- get_game_shot_allowed_report(game_id, team_id)
    game_log <- rbind(game_log, game_report)
  }
  game_log <- game_log %>% summarise(across(where(is.numeric), mean))
  return(game_log)
}


# Venue splits (allowed)
get_home_allowed_splits <- function(team_tricode, date) {
  team_id <- get_team_id(tricode = toupper(team_tricode))
  home_sched <- get_team_schedule_df(team = toupper(team_tricode), date, season = 20252026) %>%
    dplyr::filter(homeTeam.id == team_id)
  n_home <- nrow(home_sched)
  game_log <- data.frame(matrix(ncol = 4, nrow = 0))
  colnames(game_log) <- c('total_shot_attempts', 'shots_on_goal_plus_goals', 'blocked_shots', 'missed_shots')
  for (i in 1:n_home) {
    game_id <- home_sched$id[n_home - (i - 1)]
    game_report <- get_game_shot_allowed_report(game_id, team_id)
    game_log <- rbind(game_log, game_report)
  }
  home_mean_allowed <- game_log %>% summarise(across(where(is.numeric), mean))
  return(home_mean_allowed)
}

get_away_allowed_splits <- function(team_tricode, date) {
  team_id <- get_team_id(tricode = toupper(team_tricode))
  away_sched <- get_team_schedule_df(team = toupper(team_tricode), date, season = 20252026) %>%
    dplyr::filter(awayTeam.id == team_id)
  n_away <- nrow(away_sched)
  game_log <- data.frame(matrix(ncol = 4, nrow = 0))
  colnames(game_log) <- c('total_shot_attempts', 'shots_on_goal_plus_goals', 'blocked_shots', 'missed_shots')
  for (i in 1:n_away) {
    game_id <- away_sched$id[n_away - (i - 1)]
    game_report <- get_game_shot_allowed_report(game_id, team_id)
    game_log <- rbind(game_log, game_report)
  }
  away_mean_allowed <- game_log %>% summarise(across(where(is.numeric), mean))
  return(away_mean_allowed)
}

get_full_against_home_preds <- function(tricode, date){
  l8_against <- get_team_shot_allowed_log(tricode, 8, date) 
  l4_against  <- get_team_shot_allowed_log(tricode, 4, date) 
  home_against <- get_home_allowed_splits(tricode, date)
  
  l8_against <- l8_against %>%
    dplyr::select(all_of(c('total_shot_attempts', 'shots_on_goal_plus_goals', 'missed_shots'))) %>%
    rename(l8_sa_ag = total_shot_attempts, l8_sog_ag = shots_on_goal_plus_goals, l8_miss_ag = missed_shots)
  l4_against <- l4_against %>%
    dplyr::select(all_of(c('total_shot_attempts', 'shots_on_goal_plus_goals', 'missed_shots'))) %>%
    rename(l4_sa_ag = total_shot_attempts, l4_sog_ag = shots_on_goal_plus_goals, l4_miss_ag = missed_shots)
  home_against <- home_against %>%
    dplyr::select(all_of(c('total_shot_attempts', 'shots_on_goal_plus_goals', 'missed_shots'))) %>%
    rename(ha_sa_ag = total_shot_attempts, ha_sog_ag = shots_on_goal_plus_goals, ha_miss_ag = missed_shots)
  
  shot_against_preds <- cbind(l8_against, l4_against, home_against)
  return(shot_against_preds)
}


get_full_against_away_preds <- function(tricode, date){
  l8_against <- get_team_shot_allowed_log(tricode, 8, date)
  l4_against  <- get_team_shot_allowed_log(tricode, 4, date)
  away_against <- get_away_allowed_splits(tricode, date)
  
  l8_against <- l8_against %>%
    dplyr::select(all_of(c('total_shot_attempts', 'shots_on_goal_plus_goals', 'missed_shots'))) %>%
    rename(l8_sa_ag = total_shot_attempts, l8_sog_ag = shots_on_goal_plus_goals, l8_miss_ag = missed_shots)
  l4_against <- l4_against %>%
    dplyr::select(all_of(c('total_shot_attempts', 'shots_on_goal_plus_goals', 'missed_shots'))) %>%
    rename(l4_sa_ag = total_shot_attempts, l4_sog_ag = shots_on_goal_plus_goals, l4_miss_ag = missed_shots)
  away_against <- away_against %>%
    dplyr::select(all_of(c('total_shot_attempts', 'shots_on_goal_plus_goals', 'missed_shots'))) %>%
    rename(ha_sa_ag = total_shot_attempts, ha_sog_ag = shots_on_goal_plus_goals, ha_miss_ag = missed_shots)
  
  shot_against_preds <- cbind(l8_against, l4_against, away_against)
  return(shot_against_preds)
}

get_full_against_home_preds('SJS', today)
get_full_against_away_preds('SJS', '2025-11-03')

get_game_shot_allowed_report(2025020534, 'edm')
```


# Getting Days of Rest
```{r}
get_rest <- function(game_id, team_tricode, max_rest = 7) {

  sched <- get_team_schedule(team = toupper(team_tricode))

  # Defensive check
  if (!"gameDate" %in% names(sched) || !"id" %in% names(sched)) {
    return(max_rest)
  }

  sched <- sched %>%
    dplyr::mutate(gameDate = as.Date(gameDate)) %>%
    dplyr::arrange(gameDate)

  rownum <- which(sched$id == game_id)

  # Game not found (future game, preseason mismatch, etc.)
  if (length(rownum) == 0) {
    last_game_date <- max(sched$gameDate, na.rm = TRUE)
    return(min(as.numeric(Sys.Date() - last_game_date) - 1, max_rest))
  }

  # First game of season
  if (rownum == 1) {
    return(max_rest)
  }

  days_rest <- as.numeric(
    sched$gameDate[rownum] - sched$gameDate[rownum - 1]
  ) - 1

  return(min(days_rest, max_rest))
}

get_rest(2025020534, 'edm')
```



# Getting Slate Data
```{r, warning=F}
get_daily_game_reports <- function(date) {
  slate <- get_schedule(date = date) %>%
    dplyr::select(c('id', 'gameType', 'startTimeUTC', 'awayTeam.id', 'awayTeam.abbrev',
             'homeTeam.id', 'homeTeam.abbrev')) 
  
  df <- data.frame(game_id = slate[,1], team = slate[,5], opponent = slate[,7], ha = 'home')
  names(df) <- c('game_id', 'team', 'opponent', 'h/a')
  
  df2 <- data.frame(game_id = slate[,1], team = slate[,7], opponent = slate[,5], ha='away')
  names(df2) <- c('game_id', 'team', 'opponent', 'h/a')
  games <- rbind(df, df2)
  games[c('l8_sa', 'l8_sog', 'l8_miss', 'l4_sa', 'l4_sog', 'l4_miss', 'ha_sa', 'ha_sog', 'ha_miss',
          'l8_sa_ag', 'l8_sog_ag', 'l8_miss_ag', 'l4_sa_ag', 'l4_sog_ag', 'l4_miss_ag', 
          'ha_sa_ag', 'ha_sog_ag', 'ha_miss_ag', 'rest_diff', 'true_sog')] <- NA
  #return(games)
  temp4 <- 0
  for (i in 1:nrow(games)) {
    if (games$'h/a'[i] == 'home'){
      temp <- get_full_home_preds(games$team[i], date)
      common_cols <- intersect(names(games[i,]), names(temp))
      games[i, common_cols] <- temp[1, common_cols]
      
      temp2 <- get_full_against_away_preds(games$opponent[i], date)
      common_cols_2 <- intersect(names(games[i,]), names(temp2))
      games[i, common_cols_2] <- temp2[1, common_cols_2]
      
    }else if (games$'h/a'[i] == 'away'){
      temp <- get_full_away_preds(games$team[i], date)
      common_cols <- intersect(names(games[i,]), names(temp))
      games[i, common_cols] <- temp[1, common_cols]
      
      temp2 <- get_full_against_home_preds(games$opponent[i], date)
      common_cols_2 <- intersect(names(games[i,]), names(temp2))
      games[i, common_cols_2] <- temp2[1, common_cols_2]
    }
    games$rest_diff[i] <- get_rest(games$game_id[i], games$team[i]) - get_rest(games$game_id[i], games$opponent[i])
    if (date >= Sys.Date()){
      tid <- get_team_id(games$team[i])
      true_sog <- get_game_shot_report(games$game_id[i], tid)
      games$true_sog[i] <- true_sog$shots_on_goal_plus_goals
    }
  }
  return(games)
}


daily_reports <- get_daily_game_reports('2025-12-18')
#daily_reports
```


# Creating Full Training Dataset
```{r, warning=F}
start_date <- as.Date('2025-12-17')
end_date <- as.Date('2025-12-20')
train <- data.frame()

for(i in seq(start_date, end_date, by='day')){
  temp3 <- get_daily_game_reports(as.Date(i))
  #print(temp3)
  train <- rbind(train, temp3)
}
train
#train[,c('team', 'true_sog')]
wk9 <- train


train_data <- rbind(wk1, wk2, wk3, wk4, wk5, wk6, wk7, wk8, wk9)
train_data
# Start Date - 2025-10-26
# End Date - 2025-12-20

train_data$sog_trend <- train_data$l4_sog - train_data$l8_sog
train_data$sog_ag_trend <- train_data$l4_sog_ag - train_data$l8_sog_ag
train_data$home <- ifelse(train_data$'h/a' == 'home', 1, 0)
train_data$off_def_gap <- train_data$l8_sog - train_data$l8_sog_ag
train_data$pace <- train_data$l8_sog + train_data$l8_sog_ag
train_data$shot_perc <- train_data$l8_sog / train_data$l8_sa
train_data$shot_perc_ag <- train_data$l8_sog_ag / train_data$l8_sa_ag
train_data
save(train_data, file="train_data.RData")

red_train <- train_data %>% dplyr::select('game_id', 'team', 'opponent', 'l8_sog', 'l4_sog', 'ha_sog', 'l8_sog_ag', 'l4_sog_ag', 'ha_sog_ag', 'rest_diff', 'true_sog', 'sog_trend', 'sog_ag_trend', 'home', 'off_def_gap', 'pace', 'shot_perc', 'shot_perc_ag')
#head(red_train)
```

# PCA
```{r}
pca_subset <- train_data %>% dplyr::select(-c('game_id', 'team', 'opponent', 'h/a', 'l8_miss', 'l4_miss', 'ha_miss', 'l8_miss_ag', 'l4_miss_ag', 'ha_miss_ag'))

prout <- prcomp(pca_subset, scale=T)
prout$rotation
biplot(prout, scale=0)

prout$sdev
prvar <- prout$sdev^2
prvar

pve <- prvar/sum(prvar)
pve

plot(cumsum(pve), xlab="Principal Component ", 
     ylab=" Cumulative Proportion of Variance Explained ", ylim=c(0,1), type='b')
```


# Scaling
```{r}
num_vars <- c('l8_sog', 'l4_sog', 'ha_sog', 'l8_sog_ag', 'l4_sog_ag', 'ha_sog_ag', 'rest_diff', 'sog_trend', 'sog_ag_trend', 'off_def_gap', 'pace', 'shot_perc', 'shot_perc_ag')
set.seed(42)

train_idx <- createDataPartition(red_train$true_sog, p=.8, list=F)

train <- red_train[train_idx, ]
test <- red_train[-train_idx, ]

scaler <- preProcess(train[, num_vars], method=c('center', 'scale'))

train_scaled <- train
test_scaled <- test
train_scaled[, num_vars] <- predict(scaler, train[, num_vars])
test_scaled[, num_vars] <- predict(scaler, test[, num_vars])
test_scaled2 <- test_scaled
summary(train_scaled[,num_vars])

```

# Model V1
```{r}
nb_fit <- glm.nb(true_sog ~ l8_sog + sog_trend + l8_sog_ag + sog_ag_trend + rest_diff
                 + shot_perc + shot_perc_ag + home, data=train_scaled)

summary(nb_fit)
nb_fit$theta


x_train <- model.matrix(true_sog ~ l8_sog + sog_trend + l8_sog_ag + sog_ag_trend + rest_diff
                 + shot_perc + shot_perc_ag + home, data=train_scaled)[,-1]
y_train <- train_scaled$true_sog

ridge_pois <- cv.glmnet(x_train, y_train, family='poisson', alpha=0)


x_test <- model.matrix(
  true_sog ~ l8_sog + sog_trend + l8_sog_ag + sog_ag_trend +
             rest_diff + shot_perc + shot_perc_ag + home,
  test_scaled
)[, -1]

test$pred_nb    <- predict(nb_fit, test_scaled, type = "response")
test$pred_ridge <- predict(ridge_pois, x_test, s = "lambda.min", type = "response")

Metrics::rmse(test$true_sog, test$pred_nb)
Metrics::rmse(test$true_sog, test$pred_ridge)

exp(coef(nb_fit))

```


# Model V2
```{r}
re_fit <- glm.nb(
  true_sog ~ 
    l8_sog + l8_sog_ag + sog_ag_trend +
    shot_perc + shot_perc_ag + rest_diff + home +
    factor(team) + factor(opponent),
  data = train_scaled
)
summary(re_fit)
```


# Test Data (w/ team effects)
```{r}
test_scaled$pred_sog <- predict(re_fit, newdata=test_scaled, type='response')
test_scaled$pred_sog

rmse_val <- sqrt(mean((test_scaled$true_sog - test_scaled$pred_sog)^2))
mae_val  <- mean(abs(test_scaled$true_sog - test_scaled$pred_sog))

rmse_val
mae_val


baseline_pred <- mean(train_scaled$true_sog)

rmse_baseline <- rmse(test_scaled$true_sog, baseline_pred)
rmse_baseline

mean(test_scaled$pred_sog - test_scaled$true_sog)

plot(
  test_scaled$true_sog,
  test_scaled$pred_sog,
  xlab = "Actual SOG",
  ylab = "Predicted SOG"
)
abline(0, 1, col = "red")


resid <- test_scaled$true_sog - test_scaled$pred_sog

plot(
  test_scaled$pred_sog,
  resid,
  xlab = "Predicted SOG",
  ylab = "Residual"
)
abline(h = 0, col = "red")
```

# Test Data (w/o team effects)
```{r}
test_scaled2$pred_sog <- predict(nb_fit, newdata=test_scaled, type='response')
test_scaled2$pred_sog

rmse_val <- sqrt(mean((test_scaled2$true_sog - test_scaled2$pred_sog)^2))
mae_val  <- mean(abs(test_scaled2$true_sog - test_scaled2$pred_sog))
rmse_val
mae_val


baseline_pred <- mean(test_scaled2$true_sog)

rmse_baseline <- rmse(test_scaled2$true_sog, baseline_pred)
rmse_baseline

resid2 <- test_scaled2$true_sog - test_scaled2$pred_sog

plot(
  test_scaled2$pred_sog,
  resid2,
  xlab = "Predicted SOG",
  ylab = "Residual"
)
abline(h = 0, col = "red")
```


# Baseline
```{r}
baseline_pred <- rep(mean(train$true_sog), nrow(test))
rmse_baseline <- rmse(test$true_sog, baseline_pred)
mae_baseline  <- mae(test$true_sog, baseline_pred)
rmse_baseline
mae_baseline

```

# Model Save
```{r}
saveRDS(nb_fit, 'nb_sog_model.rds')
saveRDS(scaler, "sog_scaler.rds")

```


# Creating Predictions
```{r}
daily_reports <- get_daily_game_reports('2026-02-02')
daily_reports$sog_trend <- daily_reports$l4_sog - daily_reports$l8_sog
daily_reports$sog_ag_trend <- daily_reports$l4_sog_ag - daily_reports$l8_sog_ag
daily_reports$home <- ifelse(daily_reports$'h/a' == 'home', 1, 0)
daily_reports$shot_perc <- daily_reports$l8_sog / daily_reports$l8_sa
daily_reports$shot_perc_ag <- daily_reports$l8_sog_ag / daily_reports$l8_sa_ag
daily_reports$pred_sog <- NA
daily_report_red <- daily_reports %>% 
  dplyr::select(c('game_id', 'team', 'opponent', 'l8_sog', 'sog_trend', 'l8_sog_ag', 'sog_ag_trend', 'rest_diff', 'shot_perc', 'shot_perc_ag', 'home', 'pred_sog'))
daily_report_red
predictors <- c('l8_sog', 'sog_trend', 'l8_sog_ag', 'sog_ag_trend', 'rest_diff', 'shot_perc', 'shot_perc_ag')

saveRDS(df, 'daily_report_red.rds')

req_vars <- scaler$method$center
miss_vars <- setdiff(req_vars, names(daily_report_red))
daily_report_red[miss_vars] <- 0

daily_report_red_scaled <- daily_report_red
daily_report_red_scaled[req_vars] <- predict(scaler, daily_report_red[req_vars])
daily_report_red_scaled <- daily_report_red_scaled[c('game_id', 'team', 'opponent', 'l8_sog', 'sog_trend', 'l8_sog_ag', 'sog_ag_trend', 'rest_diff', 'shot_perc', 'shot_perc_ag', 'home', 'pred_sog')]
#daily_report_red_scaled

daily_report_red_scaled$pred_sog <- predict(nb_fit, newdata=daily_report_red_scaled, type='response')
daily_report_red_scaled$pred_sog <- round(daily_report_red_scaled$pred_sog, 1)
daily_report_red_scaled[c('team', 'pred_sog')]

```

# Amplifier
```{r}
eta <- predict(nb_fit, daily_report_red_scaled, type='link')
beta0 <- coef(nb_fit)[1]
effects <- eta - beta0
amp <- 2
eta_amp <- beta0 + effects * amp
daily_report_red_scaled$pred_sog <- exp(eta_amp)
daily_report_red_scaled[c('team', 'pred_sog')]
```



# Recording Results
```{r}
jan23 <- daily_report_red_scaled
saveRDS(jan23, 'jan23.rds')
```


# Amplifier + Quantiles
```{r, warning=F}
mean_sog <- mean(train_data$true_sog)
spread_factor <- 1.2
daily_report_red_scaled$pred_sog_wide <- round(mean_sog + 
  (daily_report_red_scaled$pred_sog - mean_sog) * spread_factor, 1)

mu_pred <- as.vector(daily_report_red_scaled$pred_sog)
theta <- nb_fit$theta

set.seed(42)
simulations <- matrix(NA, nrow=length(mu_pred), ncol=1000)

for(i in 1:length(mu_pred)){
  simulations[i,] <- MASS::rnegbin(1000, mu=mu_pred[i], theta=theta)
}

daily_report_red_scaled$ceiling_sog <- round(apply(simulations, 1, quantile, probs=0.75), 1)
daily_report_red_scaled$floor_sog <- round(apply(simulations, 1, quantile, probs=0.25), 1)

daily_report_red_scaled <- daily_report_red_scaled %>%
  mutate(
    team_id = sapply(team, get_team_id), # Uses your existing get_team_id function
    logo_url = paste0("https://assets.nhle.com/logos/nhl/svg/", team_id, "_light.svg")
  )

for(i in 1:nrow(daily_report_red_scaled)){
  tid <- get_team_id(daily_report_red_scaled$team[i])
  true_sog <- get_game_shot_report(daily_report_red_scaled$game_id[i], tid)
  daily_report_red_scaled$true_sog[i] <- true_sog$shots_on_goal_plus_goals
}

daily_report_red_scaled[c('logo_url', 'team', 'pred_sog', 'pred_sog_wide', 'ceiling_sog', 'floor_sog', 'true_sog')] %>%
  text_transform(
    locations = cells_body(columns = logo_url),
    fn = function(x) {
      web_image(url = x, height = 30)
    }
  )
  

ggplot(daily_report_red_scaled, aes(y=reorder(team, true_sog))) + 
  geom_segment(aes(x=true_sog, xend=pred_sog, yend=team)) + 
  geom_point(aes(x=true_sog), color='red', size=5, shape=18) + 
  geom_point(aes(x=pred_sog), color='orange', size=4) +
  geom_point(aes(x=pred_sog_wide), color='blue', size=4) + 
  geom_point(aes(x=ceiling_sog), color='green', size=2) +
  geom_point(aes(x=floor_sog), color='green', size=2) +
  theme_minimal() +
  labs(y='predicted_sog')

mean(daily_report_red_scaled$true_sog - daily_report_red_scaled$pred_sog)
mean(daily_report_red_scaled$true_sog - daily_report_red_scaled$pred_sog_wide)
median(daily_report_red_scaled$true_sog - daily_report_red_scaled$pred_sog)
median(daily_report_red_scaled$true_sog - daily_report_red_scaled$pred_sog_wide)

```

# RUN ME
```{r, warning=F}
daily_reports <- get_daily_game_reports('2026-02-02')
daily_reports$sog_trend <- daily_reports$l4_sog - daily_reports$l8_sog
daily_reports$sog_ag_trend <- daily_reports$l4_sog_ag - daily_reports$l8_sog_ag
daily_reports$home <- ifelse(daily_reports$'h/a' == 'home', 1, 0)
daily_reports$shot_perc <- daily_reports$l8_sog / daily_reports$l8_sa
daily_reports$shot_perc_ag <- daily_reports$l8_sog_ag / daily_reports$l8_sa_ag
daily_reports$pred_sog <- NA
daily_report_red <- daily_reports %>% 
  dplyr::select(c('game_id', 'team', 'opponent', 'l8_sog', 'sog_trend', 'l8_sog_ag', 'sog_ag_trend', 'rest_diff', 'shot_perc', 'shot_perc_ag', 'home', 'pred_sog'))
predictors <- c('l8_sog', 'sog_trend', 'l8_sog_ag', 'sog_ag_trend', 'rest_diff', 'shot_perc', 'shot_perc_ag')

saveRDS(df, 'daily_report_red.rds')

req_vars <- scaler$method$center
miss_vars <- setdiff(req_vars, names(daily_report_red))
daily_report_red[miss_vars] <- 0

daily_report_red_scaled <- daily_report_red
daily_report_red_scaled[req_vars] <- predict(scaler, daily_report_red[req_vars])
daily_report_red_scaled <- daily_report_red_scaled[c('game_id', 'team', 'opponent', 'l8_sog', 'sog_trend', 'l8_sog_ag', 'sog_ag_trend', 'rest_diff', 'shot_perc', 'shot_perc_ag', 'home', 'pred_sog')]

daily_report_red_scaled$pred_sog <- predict(nb_fit, newdata=daily_report_red_scaled, type='response')
daily_report_red_scaled$pred_sog <- round(daily_report_red_scaled$pred_sog, 1)


mean_sog <- mean(train_data$true_sog)
spread_factor <- 1.2
daily_report_red_scaled$pred_sog_wide <- round(mean_sog + 
  (daily_report_red_scaled$pred_sog - mean_sog) * spread_factor, 1)

mu_pred <- as.vector(daily_report_red_scaled$pred_sog)
theta <- nb_fit$theta

set.seed(42)
simulations <- matrix(NA, nrow=length(mu_pred), ncol=1000)

for(i in 1:length(mu_pred)){
  simulations[i,] <- MASS::rnegbin(1000, mu=mu_pred[i], theta=theta)
}

daily_report_red_scaled$ceiling_sog <- round(apply(simulations, 1, quantile, probs=0.75), 1)
daily_report_red_scaled$floor_sog <- round(apply(simulations, 1, quantile, probs=0.25), 1)

for(i in 1:nrow(daily_report_red_scaled)){
  tid <- get_team_id(daily_report_red_scaled$team[i])
  true_sog <- get_game_shot_report(daily_report_red_scaled$game_id[i], tid)
  daily_report_red_scaled$true_sog[i] <- true_sog$shots_on_goal_plus_goals
}


table_html <- daily_report_red_scaled %>%
  dplyr::select('team', 'opponent', 'pred_sog_wide', 'floor_sog', 'ceiling_sog', 'true_sog') %>%
  gt() %>%
  tab_header(title = paste("NHL Shot Predictions:", Sys.Date())) %>%
  cols_label(
    'floor_sog' = "25th Percentile",
    'pred_sog_wide' = "Predicted Mean",
    'ceiling_sog' = "80th Percentile"
  ) %>%
  as_raw_html()
table_html
# Save as index.html for GitHub Pages
writeLines(table_html, "index.html")
```